Normalization by Evaluation for Call-By-Push-Value and Polarized Lambda Calculus

### 2019-07-11 ppdp19/ppdp19.tex
[ ppdp19 ] switched to proceedings version

%
% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigplan,screen,fleqn,review]{acmart} % review for draft
%% DRAFT only:
\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\sloppy

@@

The terms \fbox{$\Tm N \Gamma$} of the focused lambda calculus are the
ones of CBPV minus the positive eliminations ($\tsplit$, $\tcase$,
$\tabort$), the added negative variable rule ($\nvar$), and the
necessary changes to the binders $\tabs$ and $\tbind$.
\begin{gather*}
  \nru{\nvar}{N \in \Gamma}{\Tm N \Gamma}
\qquad
  \nru{\tabs}
      {\AddHypP P {\tTm\, N} \Gamma}
      {\TmP {P \To N} \Gamma}
\\[2ex]
  \nru{\tbind}
      {\TmP {\Comp P} \Gamma \qquad \AddHypP P {\tTm\,N} \Gamma}
      {\Tm N \Gamma}
\end{gather*}
Term interpretation
\fbox{$\denty\_ : \Tm N \Gamma \to \den \Gamma \todot \den N$}
shall be as for CBPV except that we need to exchange the
interpretation function for binders
%\fbox{$
\[
\fden\_ : \Tm N {(\cext
    \Gamma P)} \to \den \Gamma \todot \den{P \To N}
.\]
%$}.
Since a binder
for $P$ performs a maximal splitting on $P$ and takes the form of a
function defined by a case (and split) tree,
applying it to a value $v$ of type $P$ amounts to a complete
matching of $v$ against the case tree and binding the remaining
atomic and negative crumbs.  This matching can be defined for a
generic evaluation function of type
$\Evvv \J \Delta \Gamma = \den \Gamma \Delta \to \J\,\Delta$.
\[
\begin{array}{l@{~}l@{~}l@{~}lcl}
  \multicolumn 6 {l} {\tmatch ~~:~~ \den P \Delta \to
    \AddHypP P {\Evv \J \Delta} \todot \Evv \J \Delta }\\
  \tmatch & x & (\phyp\,e) & \gamma & = & e\,(\gamma,\; \treturn\,x) \\
  \tmatch & b & (\nhyp\,e) & \gamma & = & e\,(\gamma,\; b) \\
  \tmatch & a & \nbranch & \gamma & = & \tmagic\;a \\
  \tmatch & (\iota_1\;a_1) & (\bbranch\;e_1\;e_2) & \gamma & = &
    \tmatch\;a_1\;e_1\;\gamma \\
  \tmatch & (\iota_2\;a_2) & (\bbranch\;e_1\;e_2) & \gamma & = &
    \tmatch\;a_2\;e_2\;\gamma \\
  \tmatch & \ttempty & (\nsplit\,e) & \gamma & = & e\,\gamma \\
  \tmatch & (a_1,a_2) & (\bsplit\,e) & \gamma & = &
\\ \multicolumn 6 {@{\qquad}l}{
    \tmatch\;a_1\;(\tmap[\Add P]\,(\tmatch\;a_2)\;e)\;\gamma
}
\\
\end{array}
\]
With instantiations $\J = \den N$ and
$\denty\_ : \Tm N \todot \Evv {\den N} \Delta$,
the interpretation $\fden t$
of binder $t : \AddHypP P {\tTm\,N} \Gamma$
is defined as follows:
\begin{multline*}
  \fdent{t}
    {(\gamma : \den \Gamma \Delta)}
    \, (\tau : \Delta \sublist \Phi)
    \, (a : \den P \Phi)
    ~~=~~
\\ \qquad
    \tmatch
      \; a
      \; (\tmap[\Add P]\;\denty\_\;t)
      \; (\tren\,\tau\,\gamma)
\end{multline*}
This completes the definition of the normalization function
$\tnorm\, (t : \Tm N \Gamma)
  = \down N \dent t {\idenv\Gamma}$.

>>>>

%
% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,screen,fleqn]{acmart} % review for draft
%% DRAFT only:
%\settopmatter{printacmref=false} % Removes citation information below abstract
%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\sloppy

@@

The terms \fbox{$\Tm N \Gamma$} of the polarized lambda calculus are the
ones of CBPV minus the positive eliminations ($\tsplit$, $\tcase$,
$\tabort$), the added negative variable rule ($\nvar$), and the
necessary changes to the binders $\tabs$ and $\tbind$.
\begin{gather*}
  \nru{\nvar}{N \in \Gamma}{\Tm N \Gamma}
\qquad
  \nru{\tabs}
      {\AddHypP P {\tTm\, N} \Gamma}
      {\TmP {P \To N} \Gamma}
\\[2ex]
  \nru{\tbind}
      {\TmP {\Comp P} \Gamma \qquad \AddHypP P {\tTm\,N} \Gamma}
      {\Tm N \Gamma}
\end{gather*}
Term interpretation
\fbox{$\denty\_ : \Tm N \Gamma \to \den \Gamma \todot \den N$}
shall be as for CBPV except that we need to exchange the
interpretation function for binders
%\fbox{$
\[
\fden\_ : \Tm N {(\cext
    \Gamma P)} \to \den \Gamma \todot \den{P \To N}
.\]
%$}.
Since a binder
for $P$ performs a maximal splitting on $P$ and takes the form of a
function defined by a case (and split) tree,
applying it to a value $v$ of type $P$ amounts to a complete
matching of $v$ against the case tree and binding the remaining
atomic and negative crumbs.  This matching can be defined for a
generic evaluation function of type
$\Evvv \J \Delta \Gamma = \den \Gamma \Delta \to \J\,\Delta$.
\[
\begin{array}{l@{~}l@{~}l@{~}lcl}
  \multicolumn 6 {l} {\tmatch ~~:~~ \den P \Delta \to
    \AddHypP P {\Evv \J \Delta} \todot \Evv \J \Delta }\\
  \tmatch & x & (\phyp\,e) & \gamma & = & e\,(\gamma,\; \treturn\,x) \\
  \tmatch & b & (\nhyp\,e) & \gamma & = & e\,(\gamma,\; b) \\
  \tmatch & a & \nbranch & \gamma & = & \tmagic\;a \\
  \tmatch & (\iota_1\;a_1) & (\bbranch\;e_1\;e_2) & \gamma & = &
    \tmatch\;a_1\;e_1\;\gamma \\
  \tmatch & (\iota_2\;a_2) & (\bbranch\;e_1\;e_2) & \gamma & = &
    \tmatch\;a_2\;e_2\;\gamma \\
  \tmatch & \ttempty & (\nsplit\,e) & \gamma & = & e\,\gamma \\
  \tmatch & (a_1,a_2) & (\bsplit\,e) & \gamma & = &
\\ \multicolumn 6 {@{\qquad}l}{
    \tmatch\;a_1\;(\tmap[\Add P]\,(\tmatch\;a_2)\;e)\;\gamma
}
\\
\end{array}
\]
With instantiations $\J = \den N$ and
$\denty\_ : \Tm N \todot \Evv {\den N} \Delta$,
the interpretation $\fden t$
of binder $t : \AddHypP P {\tTm\,N} \Gamma$
is defined as follows:
\begin{multline*}
  \fdent{t}
    {(\gamma : \den \Gamma \Delta)}
    \, (\tau : \Delta \sublist \Phi)
    \, (a : \den P \Phi)
    ~~=~~
\\ \qquad
    \tmatch
      \; a
      \; (\tmap[\Add P]\;\denty\_\;t)
      \; (\tren\,\tau\,\gamma)
\end{multline*}
This completes the definition of the normalization function
$\tnorm\, (t : \Tm N \Gamma)
  = \down N \dent t {\idenv\Gamma}$.
